---
title: "plot_MR"
author: "Rutuja"
date: "2025-04-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code adapted from [https://andrewslabucsf.github.io/MR-tutorial/]

```{r, include=FALSE}
library(data.table)
library(tidyverse)
library(TwoSampleMR)
library(LDlinkR)
library(gt)
```

Displaying and visualizing all results.

## Sporadic Miscarriage as Exposure

### Age at first birth

```{r}
mr_dat <- read.csv("exposure/mr_dat_age")
mr_res <- read.csv("exposure/mr_res_age.csv")
res_single <- read.csv("exposure/res_single_age.csv")

mr_res
```


### Chronotype

```{r}
mr_dat <- read.csv("exposure/mr_dat_chrono")
mr_res <- read.csv("exposure/mr_res_chrono.csv")
res_single <- read.csv("exposure/res_single_chrono.csv")

mr_res
```

### Income

```{r}
mr_dat <- read.csv("exposure/mr_dat_income")
mr_res <- read.csv("exposure/mr_res_income.csv")
res_single <- read.csv("exposure/res_single_income.csv")

mr_res
```

### PTSD

```{r}
mr_dat <- read.csv("exposure/mr_dat_ptsd")
mr_res <- read.csv("exposure/mr_res_ptsd.csv")
res_single <- read.csv("exposure/res_single_ptsd.csv")

mr_res
```

### Number of children ever born

```{r}
mr_dat <- read.csv("exposure/mr_dat_number")
mr_res <- read.csv("exposure/mr_res_number.csv")
res_single <- read.csv("exposure/res_single_number.csv")

mr_res
```

## Sporadic miscarriage as outcome

### Age at first birth

This relationship was found to be significant. Plotting effect size and performing sensitivity analysis. Sensitivty analysis did not flag any assumption violations.

```{r}
mr_dat <- read.csv("outcome/mr_dat_age")
mr_res <- read.csv("outcome/mr_res_age.csv")
res_single <- read.csv("outcome/res_single_age.csv")

mr_res

res_pleio <- mr_pleiotropy_test(mr_dat)

res_pleio %>%
  select(-id.exposure, -id.outcome, -outcome, -exposure) %>%
  gt() %>%
  fmt_number(
    columns = c("egger_intercept", "se")
  ) %>%
  fmt_number(
    columns = pval,
    rows = pval > 0.001,
    decimals = 3
  ) %>%
  fmt_scientific(
    columns = pval,
    rows = pval <= 0.001,
    decimals = 1
  )

# Heterogeneity statistics
res_het <- mr_heterogeneity(mr_dat, method_list = c(
  "mr_ivw_fe", "mr_ivw_mre"
))

res_het %>%
  select(-id.exposure, -id.outcome, -outcome, -exposure) %>%
  gt() %>%
  fmt_number(
    columns = Q
  ) %>%
  fmt_number(
    columns = Q_pval,
    rows = Q_pval > 0.001,
    decimals = 3
  ) %>%
  fmt_scientific(
    columns = Q_pval,
    rows = Q_pval <= 0.001,
    decimals = 1
  )

scatter_p <- mr_scatter_plot(mr_res, mr_dat)
scatter_out_p <- scatter_p[[1]] + theme_bw() +
  guides(color = guide_legend(keywidth = 0.1, label.theme = element_text(size = 10, lineheight = 0.8))) +
  theme(
    text = element_text(size = 16),
  ) +
  xlab("SNP effect on Age at First Birth") +
  ylab("SNP effect on Sporadic Miscarriages")
scatter_out_p

forrest_p <- mr_forest_plot(res_single)
forrest_p[[1]]

funnel_p <- mr_funnel_plot(res_single)
funnel_out_p <- funnel_p[[1]] + theme_bw() +
  guides(color = guide_legend(ncol = 1)) +
  theme(
    text = element_text(size = 8),
  )

funnel_out_p

res_loo <- mr_leaveoneout(mr_dat)
p3 <- mr_leaveoneout_plot(res_loo)
p3[[1]] + xlab(str_wrap("MR leave-one-out sensitivity analysis for Age at First Birth on Sporadic Miscarriages", width = 45)) +
  theme(
    axis.title = element_text(size = 16),
    text = element_text(size = 16)
  )
```

### Chronotype

```{r}
mr_dat <- read.csv("outcome/mr_dat_chrono")
mr_res <- read.csv("outcome/mr_res_chrono.csv")
res_single <- read.csv("outcome/res_single_chrono.csv")

mr_res
```

### Income

```{r}
mr_dat <- read.csv("outcome/mr_dat_income")
mr_res <- read.csv("outcome/mr_res_income.csv")
res_single <- read.csv("outcome/res_single_income.csv")

mr_res
```

### PTSD

```{r}
mr_dat <- read.csv("outcome/mr_dat_ptsd")
mr_res <- read.csv("outcome/mr_res_ptsd.csv")
res_single <- read.csv("outcome/res_single_ptsd.csv")

mr_res
```

### Number of children ever born

This relationship was found to be significant. Plotting effect size and performing sensitivity analysis. Sensitivty analysis did not flag any assumption violations.


```{r}
mr_dat <- read.csv("outcome/mr_dat_number")
mr_res <- read.csv("outcome/mr_res_number.csv")
res_single <- read.csv("outcome/res_single_number.csv")

mr_res

res_pleio <- mr_pleiotropy_test(mr_dat)

res_pleio %>%
  select(-id.exposure, -id.outcome, -outcome, -exposure) %>%
  gt() %>%
  fmt_number(
    columns = c("egger_intercept", "se")
  ) %>%
  fmt_number(
    columns = pval,
    rows = pval > 0.001,
    decimals = 3
  ) %>%
  fmt_scientific(
    columns = pval,
    rows = pval <= 0.001,
    decimals = 1
  )

# Heterogeneity statistics
res_het <- mr_heterogeneity(mr_dat, method_list = c(
  "mr_ivw_fe", "mr_ivw_mre"
))

res_het %>%
  select(-id.exposure, -id.outcome, -outcome, -exposure) %>%
  gt() %>%
  fmt_number(
    columns = Q
  ) %>%
  fmt_number(
    columns = Q_pval,
    rows = Q_pval > 0.001,
    decimals = 3
  ) %>%
  fmt_scientific(
    columns = Q_pval,
    rows = Q_pval <= 0.001,
    decimals = 1
  )

scatter_p <- mr_scatter_plot(mr_res, mr_dat)
scatter_out_p <- scatter_p[[1]] + theme_bw() +
  guides(color = guide_legend(keywidth = 0.1, label.theme = element_text(size = 10, lineheight = 0.8))) +
  theme(
    text = element_text(size = 16),
  ) +
  xlab("SNP effect on Number of Children") +
  ylab("SNP effect on Sporadic Miscarriages")
scatter_out_p

forrest_p <- mr_forest_plot(res_single)
forrest_p[[1]]

funnel_p <- mr_funnel_plot(res_single)
funnel_out_p <- funnel_p[[1]] + theme_bw() +
  guides(color = guide_legend(ncol = 1)) +
  theme(
    text = element_text(size = 8),
  )

funnel_out_p

res_loo <- mr_leaveoneout(mr_dat)
p3 <- mr_leaveoneout_plot(res_loo)
p3[[1]] + xlab(str_wrap("MR leave-one-out sensitivity analysis for Number of Children on Sporadic Miscarriages", width = 45)) +
  theme(
    text = element_text(size = 16),
  )
```
